Generating Random Logic Programs Using Constraint Programming

### 2020-10-19 talks/aiai_seminar/talk.tex
One more presentation

\author{\textbf{Paulius Dilkas}\inst{1} \and Vaishak Belle\inst{1,2}}
\title{Generating Random Logic Programs Using Constraint Programming}
\date{CP 2020}
\institute{\inst{1} University of Edinburgh, Edinburgh, UK \and \inst{2} Alan
  Turing Institute, London, UK}

@@

\begin{frame}{What Characterizes a (Probabilistic) Logic Program?}
  \begin{columns}
    \hspace*{-0.7cm}\begin{column}{0.75\textwidth}
      \begin{empheq}[left =\onslide<6->{\color{color5}\empheqlbrace}]{equation}
        \begin{align*}
      \textcolor<5->{probability}{0.2} \prob &\textcolor<2->{predicate}{\mathtt{stress}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{person}}(\textcolor<3->{variable}{\mathsf{P}}). \\
      \textcolor<5->{probability}{0.3} \prob &\textcolor<2->{predicate}{\mathtt{influences}}(\textcolor<3->{variable}{\mathsf{P}_1}, \textcolor<3->{variable}{\mathsf{P}_2}) \ifff \textcolor<2->{predicate}{\mathtt{friend}}(\textcolor<3->{variable}{\mathsf{P}_1}, \textcolor<3->{variable}{\mathsf{P}_2}). \\
      \textcolor<5->{probability}{0.1} \prob &\textcolor<2->{predicate}{\mathtt{cancer\_spont}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{person}}(\textcolor<3->{variable}{\mathsf{P}}). \\
      \textcolor<5->{probability}{0.3} \prob &\textcolor<2->{predicate}{\mathtt{cancer\_smoke}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{person}}(\textcolor<3->{variable}{\mathsf{P}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{X}}) \ifff \textcolor<2->{predicate}{\mathtt{stress}}(\textcolor<3->{variable}{\mathsf{X}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{X}}) \ifff \textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{Y}}), \textcolor<2->{predicate}{\mathtt{influences}}(\textcolor<3->{variable}{\mathsf{Y}}, \textcolor<3->{variable}{\mathsf{X}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{cancer}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{cancer\_spont}}(\textcolor<3->{variable}{\mathsf{P}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{cancer}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \tikz \coordinate (start);\textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{P}}), \textcolor<2->{predicate}{\mathtt{cancer\_smoke}}(\textcolor<3->{variable}{\mathsf{P}})\tikz \coordinate (end);. \\
                                         &\textcolor<2->{predicate}{\mathtt{person}}(\textcolor<4->{constant}{\mathit{michelle}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{person}}(\textcolor<4->{constant}{\mathit{timothy}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{friend}}(\textcolor<4->{constant}{\mathit{timothy}}, \textcolor<4->{constant}{\mathit{michelle}}).
        \end{align*}
      \end{empheq}
    \end{column}
    \begin{column}{0.25\textwidth}
      \begin{itemize}
      \item[\textcolor{predicate}{\textbullet}]<2-> predicates, arities
      \item[\textcolor{variable}{\textbullet}]<3-> variables
      \item[\textcolor{constant}{\textbullet}]<4-> constants
      \item[\textcolor{probability}{\textbullet}]<5-> probabilities
      \item[\textcolor{color5}{\textbullet}]<6-> length
      \item[\textcolor{color6}{\textbullet}]<7-> complexity
      \end{itemize}
    \end{column}
  \end{columns}
  \onslide<7->{
    \begin{tikzpicture}[overlay]
      \draw [decorate,decoration={brace,amplitude=10pt,mirror},color=color6] (start) -- (end);
    \end{tikzpicture}
  }
\end{frame}

@@

\begin{frame}{Scalability}
  \centering
  \input{impact.tex}
\end{frame}

@@

\begin{frame}{Inference and Knowledge Compilation}
  \begin{description}
  \item[NNF] negation normal form
  \item[d-DNNF] deterministic decomposable negation normal form
  \item[BDD] binary decision diagrams
  \item[SDD] sentential decision diagrams
  \item[$k$-Best] only use the \alert{$k$} most probable proofs
%    \begin{itemize}
%    \item<2-> for every pair \alert{$\alpha \lor \beta$}, we have \alert{$\alpha
%        \land \beta = \bot$}
%    \item<2-> for every pair \alert{$\alpha \land \beta$}, no atoms are shared
%      between \alert{$\alpha$} and \alert{$\beta$}
%    \end{itemize}
  \end{description}
\end{frame}

@@

\begin{frame}{Properties of Programs vs. Inference Algorithms}
  \centering
  \input{../paper/bars.tex}
\end{frame}

\section{Conclusions}

\begin{frame}{Conclusions and Future Work}
  \begin{itemize}
  \item The model can generate (approximately) realistic instances of reasonable
    size
  \item The main performance bottleneck can be addressed by generating programs
    with a simpler structure
  \item Open questions for future work
    \begin{itemize}
    \item Can the model be used to ensure uniform sampling?
    \item What is the reason behind all algorithms behaving similarly?
    \item Why does independence have no effect on inference time?
%    \item Can random program generation be useful in learning?
    \end{itemize}
  \end{itemize}
  \begin{block}{The implementation of the model is available at}
    \url{https://github.com/dilkas/random-logic-programs}
  \end{block}
\end{frame}

\end{document}

>>>>

\author{\textbf{Paulius Dilkas} \and Vaishak Belle}
\title{Generating Random Logic Programs Using Constraint Programming}
\date{AIAI Seminar}

@@

\begin{frame}{What Characterises a (Probabilistic) Logic Program?}
  \begin{columns}
    \hspace*{-0.7cm}\begin{column}{0.75\textwidth}
      \begin{empheq}[left =\onslide<6->{\color{color5}\empheqlbrace}]{equation}
        \begin{align*}
      \textcolor<5->{probability}{0.2} \prob &\textcolor<2->{predicate}{\mathtt{stress}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{person}}(\textcolor<3->{variable}{\mathsf{P}}). \\
      \textcolor<5->{probability}{0.3} \prob &\textcolor<2->{predicate}{\mathtt{influences}}(\textcolor<3->{variable}{\mathsf{P}_1}, \textcolor<3->{variable}{\mathsf{P}_2}) \ifff \textcolor<2->{predicate}{\mathtt{friend}}(\textcolor<3->{variable}{\mathsf{P}_1}, \textcolor<3->{variable}{\mathsf{P}_2}). \\
      \textcolor<5->{probability}{0.1} \prob &\textcolor<2->{predicate}{\mathtt{cancer\_spont}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{person}}(\textcolor<3->{variable}{\mathsf{P}}). \\
      \textcolor<5->{probability}{0.3} \prob &\textcolor<2->{predicate}{\mathtt{cancer\_smoke}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{person}}(\textcolor<3->{variable}{\mathsf{P}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{X}}) \ifff \textcolor<2->{predicate}{\mathtt{stress}}(\textcolor<3->{variable}{\mathsf{X}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{X}}) \ifff \textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{Y}}), \textcolor<2->{predicate}{\mathtt{influences}}(\textcolor<3->{variable}{\mathsf{Y}}, \textcolor<3->{variable}{\mathsf{X}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{cancer}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{cancer\_spont}}(\textcolor<3->{variable}{\mathsf{P}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{cancer}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \tikz \coordinate (start);\textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{P}}), \textcolor<2->{predicate}{\mathtt{cancer\_smoke}}(\textcolor<3->{variable}{\mathsf{P}})\tikz \coordinate (end);. \\
                                         &\textcolor<2->{predicate}{\mathtt{person}}(\textcolor<4->{constant}{\mathit{michelle}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{person}}(\textcolor<4->{constant}{\mathit{timothy}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{friend}}(\textcolor<4->{constant}{\mathit{timothy}}, \textcolor<4->{constant}{\mathit{michelle}}).
        \end{align*}
      \end{empheq}
    \end{column}
    \begin{column}{0.25\textwidth}
      \begin{itemize}
      \item[\textcolor{predicate}{\textbullet}]<2-> predicates, arities
      \item[\textcolor{variable}{\textbullet}]<3-> variables
      \item[\textcolor{constant}{\textbullet}]<4-> constants
      \item[\textcolor{probability}{\textbullet}]<5-> probabilities
      \item[\textcolor{color5}{\textbullet}]<6-> length
      \item[\textcolor{color6}{\textbullet}]<7-> complexity
      \end{itemize}
    \end{column}
  \end{columns}
  \onslide<7->{
    \begin{tikzpicture}[overlay]
      \draw [decorate,decoration={brace,amplitude=10pt,mirror},color=color6] (start) -- (end);
    \end{tikzpicture}
  }
\end{frame}

@@

@@

\begin{frame}{Inference and Knowledge Compilation}
  \begin{description}
  \item[NNF] negation normal form
  \item[d-DNNF] deterministic decomposable negation normal form
  \item[BDD] binary decision diagrams
  \item[SDD] sentential decision diagrams
  \end{description}
\end{frame}

@@

%\begin{frame}{Properties of Programs vs. Inference Algorithms}
%  \centering
%  \input{../../paper/bars.tex}
%\end{frame}

\section{Summary}

\begin{frame}{Summary}
  \begin{itemize}
  \item The model can generate (approximately) realistic instances of reasonable
    size.
  \item Open questions for future work:
    \begin{itemize}
    \item Can we ensure uniform sampling?
    \item Why do all of the algorithms behave so similarly?
    \item Why does independence have no effect on inference time?
    \end{itemize}
  \end{itemize}
\end{frame}
\end{document}
